My title

### 2024-03-15 paper/paper.qmd
Added to Quarto page with some graphs

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

@@

cleaned_data <- read_csv("/Users/liamwall/Ontario-CA-Mortality/data/analysis_data/cleaned_data.csv")

#Total annual death data for both sexes and all causes
total_yearly_death_data <- cleaned_data |>
  filter(characteristics == "Number of deaths",
         cause == "Total, all causes of death [A00-Y89]",
         age == "Age at time of death, all ages",
         sex == "Both sexes") |>
  mutate(number_of_deaths = value) |>
  select(year, age, sex, cause, number_of_deaths) |>
  ggplot(aes(x = year, y = number_of_deaths)) +
  geom_point()

@@

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jacob Gilbert
  - Liam Wall
thanks: "Code and data are available at: https://github.com/JfpGilbert0/Ontario-CA-Mortality."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: fig-all_values
#| fig-cap: "Visual of each value in the dataset, the number of deaths per year per cause. Each color corresponds to a cause"
#| eval: true
#| echo: false
#| message: false
#| warning: false

cleaned_data <- read_csv("/Users/liamwall/Ontario-CA-Mortality/data/analysis_data/cleaned_data.csv")
joint_data <- read_csv(("/Users/liamwall/Ontario-CA-Mortality/data/analysis_data/joint_data.csv"))

cleaned_data |>
  filter(characteristics == "Number of deaths",
         cause != "Total, all causes of death [A00-Y89]",
         age == "Age at time of death, all ages",
         sex == "Both sexes") |>
  mutate(number_of_deaths = value) |>
  select(year, age, sex, cause, number_of_deaths) |>
  ggplot(aes(x = year, y = number_of_deaths)) +
  geom_point(aes(color = cause), alpha = 0.5) +
  theme(legend.position = "none") +
  labs(x = "Year", y = "Total Deaths", title = "Top 43 Causes of Death in Ontario")

joint_data |>
  ggplot(aes(x = year, y = value)) +
  geom_point(aes(color = cause), alpha = 0.5) +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Total Deaths", title = "Top 10 Causes of Death in Ontario")

```

```{r}
#| label: fig-top_three_bottom_seven
#| fig-cap: "Visual of the trends in top three causes versus the bottom seven."
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| layout-ncol: 2

joint_data |>
  filter(year_rank < 4) |>
  ggplot(aes(x = year, y = value)) +
  geom_point(aes(color = cause), alpha = 0.5) +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Total Deaths", title = "Top Three Causes of Death in Ontario")

joint_data |>
  filter(year_rank > 3) |>
  ggplot(aes(x = year, y = value)) +
  geom_point(aes(color = cause), alpha = 0.5) +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Total Deaths", title = "Top 10 to 7 Causes of Death in Ontario")

```

@@

#Total annual death data for both sexes and all causes
total_yearly_death_data <- cleaned_data |>
  filter(characteristics == "Number of deaths",
         cause == "Total, all causes of death [A00-Y89]",
         age == "Age at time of death, all ages",
         sex == "Both sexes") |>
  mutate(number_of_deaths = value) |>
  select(year, age, sex, cause, number_of_deaths) |>
  ggplot(aes(x = year, y = number_of_deaths)) +
  geom_point() +
  labs(x = "Year", y = "Total Deaths", title = "Annual Deaths in Ontario")

@@

```{r}
#| label: fig-model-one
#| fig-cap: "Model regressing total deaths based on year."
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| layout-ncol: 2

# could also add a predictor which would be year or year_rank
cause_of_death_ontario_neg_binomial <-
  stan_glm(
    value ~ cause,
    data = joint_data,
    family = neg_binomial_2(link = "log"),
    seed = 853
  )

summary(cause_of_death_ontario_neg_binomial)

pp_check(cause_of_death_ontario_neg_binomial)

```
